systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2005-11-07 paper/trunk/systemfit.tex
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@584 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D. Hamann} %% comma-separated
\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D. Hamann\\
   Forest Informatics, Inc.\\
   PO Box 1421\\
   Corvallis, Oregon 97339-1421\\
   e-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

@@

% \section{CVS Revision Log}
% \begin{verbatim}
% $Log$
% Revision 1.16  2005/11/02 18:31:45  hamannj
% small last changes
%
% Revision 1.15  2005/11/02 18:15:40  hamannj
% small changes

>>>>

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Arne Henningsen, Jeff D.\ Hamann} %% comma-separated
\Plaintitle{systemfit: A Package to Estimate Simultaneous Equation Systems 
            in R} %% without formatting
\Shorttitle{systemfit} %% a short title (if necessary)

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
   Arne Henningsen\\
   Department of Agricultural Economics\\
   University of Kiel\\
   D-24098 Kiel, Germany\\
   E-mail: \email{ahenningsen@agric-econ.uni-kiel.de}\\
   URL: \url{http://www.uni-kiel.de/agrarpol/ahenningsen/}\\
   \\
   Jeff D.\ Hamann\\
   Forest Informatics, Inc.\\
   PO Box 1421\\
   Corvallis, Oregon 97339-1421\\
   e-mail: \email{jeff.hamann@forestinformatics.com}\\
   URL: \url{http://www.forestinformatics.com}\\
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

@@

% \section{CVS Revision Log}
% \begin{verbatim}
% $Log$
% Revision 1.18  2005/11/07 14:08:56  henningsena
% a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
%
% Revision 1.17  2005/11/07 10:21:59  henningsena
% just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
%
% Revision 1.16  2005/11/02 18:31:45  hamannj
% small last changes
%
% Revision 1.15  2005/11/02 18:15:40  hamannj
% small changes

### 2005-11-07 paper/trunk/systemfit_intro.tex
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@584 591db2bd-e0a0-4d22-8c19-2192dbb7da31

Many theoretical models consist of more than one equation. If these
models are econometrically estimated, it may not be appropriate to
estimate the equations individually, but all together, because the
disturbance terms of these equations are likely to be
contemporaneously correlated \citep{theil71} (make sure this is the
correct paper). Estimating all equations simultaneously, taking the
covariance structure of the residuals into account, leads to efficient
estimates. Ignoring this correlation leads to an inefficient parameter
estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation approach
especially the economic theory suggests many cross-equation
restrictions \textbf{(is this statement really true?)}.

The \pkg{systemfit} package provides the capability to estimate linear
and non-linear equation systems in \proglang{R}.  Although linear and
non-linear equation systems can be estimated with several other
statistical and econometric software packages (e.g. \proglang{SAS},
\proglang{EViews}, \proglang{TSP}), \pkg{systemfit} has several
advantages.  First, all estimation procedures are publicly available
in the source code.  Second, the estimation algorithms can be easily
modified to meet specific requirements.  Third, the (advanced) user
can control many important details of the estimation procedure that
are generally not available in other software packages.  On the other
hand, these user options have reasonable defaults that also beginners
can easily use the provided functions.

>>>>

Many theoretical models consist of more than one equation. If these
models are econometrically estimated, it may not be appropriate to
estimate the equations individually, but all together, because the
disturbance terms of these equations are likely to be contemporaneously
correlated. %\citep{theil71} (make sure this is the correct paper).
Estimating all equations simultaneously, taking the
covariance structure of the residuals into account, leads to efficient
estimates.
Ignoring this correlation leads to an inefficient
parameter estimation \citep{zellner62}.

Another reason to estimate an equation system simultaneously are
cross-equation parameter restrictions. These restrictions can be
tested and/or imposed only in a simultaneous estimation
approach especially the economic theory suggests many cross-equation
restrictions.

The \pkg{systemfit} package provides the capability to estimate
linear and non-linear equation systems in \proglang{R}.
Although linear and non-linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control many important details of the
estimation procedure that are generally not available in other software
packages.
On the other hand, these user options have reasonable defaults that also
beginners can easily use the provided functions.

### 2005-11-07 paper/trunk/systemfit_other.tex
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of a sentence)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@584 591db2bd-e0a0-4d22-8c19-2192dbb7da31

However, this yields an unbiased estimator only if $K_i = K_j$ 
\citep[p. 469]{judge85}. 
% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
% whereas the first is unbiased only if i = j. 
% However, if K_i = K_j the first and the second are equal.
% Why is the first biased if K_i = K_j ???????????

A further approach to obtain the estimated residual variance-covariance
matrix is \citep[p. 309]{zellner62c}
\begin{eqnarray}
   \widehat{\sigma}_{ij} & = & 
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
   & = &
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
   \label{eq:rcov2}
\end{eqnarray} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.322]{theil71}. 
Hence, °it is doubtful whether [this formula] is really superior to 
[(\ref{eq:rcov0})]° \citep[p.322]{theil71}.

The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
if the estimated residual variance-covariance matrix is calculated
using the residuals from a first step OLS or 2SLS estimation.
It is also possible to iterate the estimation procedure and 
calculate the residual variance-covariance matrix from the
residuals of the previous iteration. 
If equation (\ref{eq:rcov0}) is applied to calculate the estimated
residual variance-covariance matrix, the results are maximum
likelihood \citep[p. 345]{greene02}.

@@

The goodness of fit of the whole system can be measured by the
McElroy's $R^2$ value \citep{mcelroy77}: 
% also: \citep[p. 345]{greene02}
\begin{equation}
   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
   { y' \left( \widehat{ \Sigma }^{-1} \otimes
   \left( I - \frac{i i'}{T} \right) \right) y }
\end{equation}
where $T$ is the number of observations in each equation,
$I$ is an $T \times T$ identity matrix and 
$i$ is a column vector of $T$ ones.

>>>>

However, this yields an unbiased estimator only if $K_i = K_j$ 
\citep[p.\ 469]{judge85}. 
% Greene (2002, p. 344) says that the second is unbiased if i = j or K_i = K_j,
% whereas the first is unbiased only if i = j. 
% However, if K_i = K_j the first and the second are equal.
% Why is the first biased if K_i = K_j ???????????

A further approach to obtain the estimated residual variance-covariance
matrix is \citep[p.\ 309]{zellner62c}
\begin{eqnarray}
   \widehat{\sigma}_{ij} & = & 
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ X_i \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' \right] } \\
   & = &
   \frac{ \widehat{u}_i' \widehat{u}_j } 
   { T - K_i - K_j + tr \left[ \left( X_i' X_i \right)^{-1}
   X_i' X_j \left( X_j' X_j \right)^{-1} X_j' X_i \right] } \\
   \label{eq:rcov2}
\end{eqnarray} 
This yields an unbiased estimator for all elements of $\widehat{\Sigma}$,
but even if $\widehat{\Sigma}$ is an unbiased estimator of $\Sigma$, 
its inverse $\widehat{\Sigma}^{-1}$ is not an unbiased estimator 
of $\Sigma^{-1}$ \citep[p.\ 322]{theil71}.
Furthermore, the variance-covariance matrix calculated by (\ref{eq:rcov2})
is not necessarily positive semidefinite \citep[p.\ 322]{theil71}. 
Hence, °it is doubtful whether [this formula] is really superior to 
[(\ref{eq:rcov0})]° \citep[p.\ 322]{theil71}.

The WLS, SUR, W2SLS and 3SLS parameter estimates are consistent,
if the estimated residual variance-covariance matrix is calculated
using the residuals from a first step OLS or 2SLS estimation.
It is also possible to iterate the estimation procedure and 
calculate the residual variance-covariance matrix from the
residuals of the previous iteration. 
If equation (\ref{eq:rcov0}) is applied to calculate the estimated
residual variance-covariance matrix, the results are maximum
likelihood \citep[p.\ 345]{greene02}.

@@

The goodness of fit of the whole system can be measured by the
McElroy's $R^2$ value \citep{mcelroy77}: 
% also: \citep[p.\ 345]{greene02}
\begin{equation}
   R_*^2 = 1 - \frac{ \widehat{u}' \widehat{ \Omega }^{-1} \widehat{u} }
   { y' \left( \widehat{ \Sigma }^{-1} \otimes
   \left( I - \frac{i i'}{T} \right) \right) y }
\end{equation}
where $T$ is the number of observations in each equation,
$I$ is an $T \times T$ identity matrix and 
$i$ is a column vector of $T$ ones.

### 2005-11-07 paper/trunk/systemfit_restrictions.tex
small changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@578 591db2bd-e0a0-4d22-8c19-2192dbb7da31
small last changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@579 591db2bd-e0a0-4d22-8c19-2192dbb7da31
just changed some line breaks (deleted new ones and restored deleted ones) to make it easier to find the changes that Jeff committed in the last revision(s)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@582 591db2bd-e0a0-4d22-8c19-2192dbb7da31
a few minor changes (predominantly adding a "\ " after a period (.) that does not indicate an end of

To impose these restrictions the $X$ matrix is
(post-)multiplied by this $T$ matrix.
\begin{equation}
    X^* = X \cdot T
\end{equation}

@@

The second way to impose linear parameter restrictions 
can be formulated by
\begin{equation}
   R \beta^0 = q
\end{equation}
where $\beta^0$ is the vector of the restricted coefficients, 
$R$ is a matrix and $q$ is a vector (see \citealp[p. 100]{greene02}). 
Each restriction belongs to one row of $R$ and the corresponding 
element of~$q$.

The first way is less flexible than this latter one%
\footnote{e.g.\ restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
by the first method.}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g. (near) singularity of the matrices to be inverted),
while the first doesn't.

@@

(***** This has to be checked!!!!! **************)

>>>>

To impose these restrictions, the $X$ matrix is
(post-)multiplied by this $T$ matrix.
\begin{equation}
    X^* = X \cdot T
\end{equation}

@@

The second way to impose linear parameter restrictions 
can be formulated by
\begin{equation}
   R \beta^0 = q
\end{equation}
where $\beta^0$ is the vector of the restricted coefficients, 
$R$ is a matrix and $q$ is a vector (see \citealp[p.\ 100]{greene02}). 
Each restriction belongs to one row of $R$ and the corresponding 
element of~$q$.

The first way is less flexible than this latter one%
\footnote{e.g.\ restrictions like $\beta_1 + \beta_2 = 4$ cannot be imposed
by the first method.}, 
but the first way is preferable if equality constraints for coefficients
across many equations of the system are imposed. 
Of course, these restrictions can be also imposed using
the latter method.
However, while the latter method increases the dimension of the 
matrices to be inverted during estimation, the first reduces it. 
Thus, in some cases the latter way leads to estimation problems
(e.g.\ (near) singularity of the matrices to be inverted),
while the first doesn't.

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "systemfit"
%%% End:
